Climate indicators

### 2023-02-28 DRAFT_climate_indicators.Rmd
import julian dates

```{r}
x_aug <- x_sub %>%
  filter(time %within% myInterval)
st_get_dimension_values(x_aug, "time")
```

@@

```{r}
x_summerMean <-   x_aug %>%
  st_apply(1:2, mean) %>%
  setNames(paste0("v_",temp_names))
```

@@

# Setting up the parameters
dir <- substr(getwd(), 1,2)
path <- ifelse(dir == "C:",
       "P:/", 
       "/data/P-Prosjekter2/")
path2 <- paste0(path, 
                "41201785_okologisk_tilstand_2022_2023/data/climate_indicators/dummy_data")
start_month <-  "06"
end_month <- "08"
myFiles <- list.files(path2, full.names = T)
summerTemp_fullSeries <- NULL

@@

  year_temp <- year(st_get_dimension_values(temp, "time")[1])
  start <- ym(paste(year_temp, start_month, sep="-"))
  end <-  ym(paste(year_temp, end_month, sep="-"))
  myInterval <- interval(start, end)

  temp <- temp %>%
    filter(time %within% myInterval) %>%
    st_apply(1:2, mean) %>%
    setNames(paste0("v_",year_temp))

@@

Now let's try it on the real data

```{r}
path <- path <- ifelse(dir == "C:", 
      "R:/",
      "/data/R/")

path2 <- paste0(path, "GeoSpatialData/Meteorology/Norway_SeNorge2018_v22.09/Original/rr_tm/") #seNorge2018_1957.nc

myFiles <- list.files(path2, full.names = T)[1:2]
start_month <-  "06"
end_month <- "08"
real_temp_summer <- NULL

@@

  tic("init")
  temp <- stars::read_ncdf(paste(path2, var="tg"))
  year_temp <- year(st_get_dimension_values(temp, "time")[1])
  start <- ym(paste(year_temp, start_month, sep="-"))
  end <-  ym(paste(year_temp, end_month, sep="-"))
  myInterval <- interval(start, end)
  toc()

  tic("middle")
  temp <- temp %>%
    filter(time %within% myInterval) %>%
    st_apply(1:2, mean) %>%
    setNames(paste0("v_",year_temp))
  toc()

  tic("c")
  real_temp_summer <- c(temp, real_temp_summer)
  rm(temp)
  toc()
}
```

>>>>

```{r}
x_aug <- x %>%
  select(Attribute_A) %>%
  adrop() %>%
  filter(time %within% myInterval)
st_get_dimension_values(x_aug, "time")
```

@@

```{r}
# setNams dont work on stars proxy obejct, so I use rename instead
lookup <- setNames("mean", paste0("v_", temp_names))

x_summerMean <-   x_aug %>%
  st_apply(1:2, mean) %>%
  rename(all_of(lookup))
```

@@

# Setting up the parameters
dir <- substr(getwd(), 1,2)
path <- ifelse(dir == "C:",
       "P:/", 
       "/data/P-Prosjekter2/")
path2 <- paste0(path, 
                "41201785_okologisk_tilstand_2022_2023/data/climate_indicators/dummy_data")
start_month <-  "06"
end_month <- "08"
myFiles <- list.files(path2, pattern=".rds", full.names = T)
summerTemp_fullSeries <- NULL

@@

  year_temp <- year(st_get_dimension_values(temp, "time")[1])
  start <- ym(paste(year_temp, start_month, sep="-"))
  end <-  ym(paste(year_temp, end_month, sep="-"))
  myInterval <- interval(start, end)
  lookup <- setNames("mean", paste0("v_", year_temp))

  temp <- temp %>%
    filter(time %within% myInterval) %>%
    st_apply(1:2, mean) %>%
    rename(all_of(lookup))

@@

Now let's try it on the real data.

_I initially tried to read stars proxy object and add steps to the call_list and execute these with `st_as_stars`, but this still returned proxy objects. Therefore I read the whole file in without proxy. To save time I can subset based on Julian dates, but this is a hard coded solution._

```{r, warning=F}
path <- path <- ifelse(dir == "C:", 
      "R:/",
      "/data/R/")

path2 <- paste0(path, "GeoSpatialData/Meteorology/Norway_SeNorge2018_v22.09/Original/rr_tm/")

myFiles <- list.files(path2, pattern=".nc$",full.names = T)[1:2]
start_month <-  "06"
end_month <- "08"
real_temp_summer <- NULL

# Get julian days
temp <- stars::read_ncdf(paste(myFiles[1]))
start_month_num <-  6
end_month_num <- 8

julian_start <- yday(st_get_dimension_values(temp, "time")[1] %m+% months(+start_month_num))
julian_end <- yday(st_get_dimension_values(temp, "time")[1] %m+% months(+end_month_num))
step <- julian_end-julian_start

@@

  tic("init")
  temp <- stars::read_ncdf(paste(myFiles[i]), var="tg", proxy=F,
                           ncsub = cbind(start = c(1, 1, julian_start), 
                              count = c(NA, NA, step)))
  year_temp <- year(st_get_dimension_values(temp, "time")[1])
  start <- ym(paste(year_temp, start_month, sep="-"))
  end <-  ym(paste(year_temp, end_month, sep="-"))
  #myInterval <- interval(start, end)
  myInterval <- interval(start, start+1)
  lookup <- setNames("mean", paste0("v_", year_temp))
  st_crs(temp) <- 25833
  toc()

  tic("middle")
  temp <- temp %>%
    filter(time %within% myInterval) %>%
    st_apply(1:2, mean) %>%
    rename(all_of(lookup)) 
  toc()

  tic("c")
  real_temp_summer <- c(temp, real_temp_summer)
  #rm(temp)
  toc()
}

```{r}
tif = system.file("tif/L7_ETMs.tif", package = "stars")
t1 = as.Date("1970-05-31")
restProxy = read_stars(tif,
               proxy = T)
class(restProxy)
restProxy2 <- st_as_stars(restProxy)
class(restProxy2)
dim(restProxy)

restProxy3 <- restProxy %>% filter(band==1)
dim(restProxy3)
restProxy4 <- st_as_stars(restProxy3)
dim(restProxy4)

restProxy5 <- restProxy %>% filter(band==1) %>%
  st_apply(1:2, mean)
dim(restProxy5)
restProxy6 <- st_as_stars(restProxy5)
dim(restProxy6)

restProxy7 <- restProxy %>% filter(band==1) 
restProxy7 <- restProxy7 %>%  st_apply(1:2, mean)
dim(restProxy7)
restProxy8 <- st_as_stars(restProxy7)
dim(restProxy8)

```{r}
dim(temp)
st_get_dimension_values(temp, "time")
#
#tic("execute")
#
real_temp_summer2 <- st_as_stars(real_temp_summer)
real_temp_summer <- st_as_stars(real_temp_summer)

#toc()
temp <- st_as_stars(temp)

write_stars(temp, "P:/41201785_okologisk_tilstand_2022_2023/data/climate_indicators/dummy_data/temp_test.tif")

write_stars(real_temp_summer, "P:/41201785_okologisk_tilstand_2022_2023/data/climate_indicators/dummy_data/real_temp_summer.tif")

dim(temp)

dim(real_temp_summer)
names(real_temp_summer)

st_get_dimension_values(real_temp_summer, "time")
```
